08-iris_comparison

### 2019-03-12 journals/08-iris_comparison/08-iris_comparison.Rmd
finished iris random forest comparison journal (08)

```{r}
# Load libraries
library(caret)
library(furrr)
library(future)
library(gretchenalbrecht)

@@

@@

# Randomly select two cases from within each of the three species of irises
selected <- c(sample(1:50, 4), sample(51:100, 4), sample(101:151, 4))

# Split up the features of the data into training and testing parts
iris_train <- iris[-selected, ]
iris_test <- iris[selected, ] %>% mutate(case = selected)
```

@@

```{r}
# Random forest model run on the iris data
iris_model <- train(iris_train %>% select(-Species), iris_train %>% pull(Species), method = 'rf')

# Predictions made using the random forest model on the testing data
iris_model_predict <- predict(iris_model, iris_test)

# Matrix of observed and predicted values
iris_test_obs_pred <- matrix(c(iris_test %>% pull(Species), iris_model_predict), ncol = 2, byrow = FALSE)

# Specify the column names
colnames(iris_test_obs_pred) <- c("Observed", "Predicted")

@@

@@

@@

  # Join the data and the explanations and edit and add additional variables
  iris_test_explain <- iris_test %>%
    mutate(case = as.character(case),
           rfscore = predict(iris_model, iris_test, type = "prob") %>% pull(virginica)) %>%
    full_join(iris_explain, by = "case") %>%
    mutate(case = factor(case),
           feature_desc = factor(feature_desc),
           feature = factor(feature),
           nbins = factor(nbins)) %>%
    arrange(nbins) %>%
    mutate(nbins = as.numeric(as.character(nbins)),
           situation = ifelse(bin_continuous == TRUE & bin_method == "quantile_bins", 
                              sprintf("%.0f quantile", nbins),
                              ifelse(bin_continuous == TRUE & bin_method == "equally_spaced",
                                     sprintf("%.0f equally spaced", nbins),
                                     ifelse(bin_continuous == TRUE & bin_method == "tree" &
                                              response == "Species",
                                            sprintf("%.0f tree", nbins),
                                            ifelse(bin_continuous == FALSE & 
                                                   use_density == TRUE, 
                                                   "kernel density", 
                                                   "normal approximation")))) %>%
             fct_relevel("2 quantile", "3 quantile", "4 quantile",
                         "5 quantile", "6 quantile", "2 equally spaced",
                         "3 equally spaced", "4 equally spaced",
                         "5 equally spaced", "6 equally spaced",
                         "2 Species tree", "3 Species tree",
                         "4 Species tree", "5 Species tree",
                         "6 Species tree")) %>%
    mutate(bin_situation = ifelse(bin_method == "quantile_bins" & 
                                  bin_continuous == TRUE,
                                  "quantile",
                                  ifelse(bin_method == "equally_spaced" & 
                                         bin_continuous == TRUE,
                                         "equally spaced", 
                                         ifelse(bin_method == "tree" & 
                                                bin_continuous == TRUE & 
                                                response == "Species",
                                                "Species tree", 
                                                ifelse(bin_continuous == FALSE & 
                                                       use_density == TRUE, 
                                                       "kernel density", 
                                                       "normal approximation"))))) %>%
    mutate(bin_situation = factor(bin_situation))

@@

This code creates and saves a dataset with summaries of the explanations.

@@

@@

@@

@@

@@

@@

>>>>

```{r}
# Load libraries
library(caret)
library(furrr)
library(future)
library(gretchenalbrecht)
library(lime)

@@

Iris is randomly split into training and testing datasets such that all of the species are represented in the testing data.

@@

# Randomly select two cases from within each of the three species of irises
selected <- c(sample(1:50, 4), sample(51:100, 4), sample(101:150, 4))

# Determine the case numbers that were not selected
cases <- 1:150
not_selected <- cases[!(cases %in% selected)]

# Split up the features of the data into training and testing parts
iris_train <- iris[-selected, ]
iris_test <- iris[selected, ] %>% mutate(case = 1:n())
```

@@

A random forest model is fit to the iris data. The predictions from the model for the testing data are shown in the table below with the actual observed values. The model gets all of the predictions correct.

```{r}
# Random forest model run on the iris training data
iris_model <- train(x = iris_train %>% select(-Species), 
                    y = iris_train %>% pull(Species), 
                    method = 'rf')

# Predictions made using the random forest model on the testing data
iris_model_predict <- predict(iris_model, iris_test %>% select(-Species, - case))

# Matrix of observed and predicted values
iris_test_obs_pred <- data.frame(Observed = iris_test %>% pull(Species),
                                 Predicted = iris_model_predict)

@@

LIME is applied using all density estimation methods with 2 to 5 bins for all of the bin based methods.

@@

The explanations from LIME are joined with the testing data. Additionally, the random forest scores for the probability that a case is species virginica are added to the dataset. This is the only species that will be considered (for now at least).

@@

  # Join the data and the explanations and edit and add additional variables
  iris_test_explain <- iris_test %>%
    mutate(case = as.character(case),
           rfscore = predict(iris_model, iris_test, type = "prob") %>% pull(virginica)) %>%
    full_join(iris_explain, by = "case") %>%
    mutate(case = factor(case),
           feature = factor(feature),
           nbins = factor(nbins)) %>%
    arrange(nbins) %>%
    mutate(nbins = as.numeric(as.character(nbins)),
           situation = ifelse(bin_continuous == TRUE & bin_method == "quantile_bins", 
                              sprintf("%.0f quantile", nbins),
                              ifelse(bin_continuous == TRUE & bin_method == "equally_spaced",
                                     sprintf("%.0f equally spaced", nbins),
                                     ifelse(bin_continuous == TRUE & bin_method == "tree" &
                                              response == "Species",
                                            sprintf("%.0f tree", nbins),
                                            ifelse(bin_continuous == FALSE & 
                                                   use_density == TRUE, 
                                                   "kernel density", 
                                                   "normal approximation")))) %>%
             fct_relevel("2 quantile", "3 quantile", "4 quantile",
                         "5 quantile", "6 quantile", "2 equally spaced",
                         "3 equally spaced", "4 equally spaced",
                         "5 equally spaced", "6 equally spaced",
                         "2 tree", "3 tree",
                         "4 tree", "5 tree",
                         "6 tree")) %>%
    mutate(bin_situation = ifelse(bin_method == "quantile_bins" & 
                                  bin_continuous == TRUE,
                                  "quantile",
                                  ifelse(bin_method == "equally_spaced" & 
                                         bin_continuous == TRUE,
                                         "equally spaced", 
                                         ifelse(bin_method == "tree" & 
                                                bin_continuous == TRUE & 
                                                response == "Species",
                                                "tree", 
                                                ifelse(bin_continuous == FALSE & 
                                                       use_density == TRUE, 
                                                       "kernel density", 
                                                       "normal approximation"))))) %>%
    mutate(bin_situation = factor(bin_situation))

@@

Summaries of the explanations are computed and saved in a dataset.

@@

All of the results in this section are in terms of the species virginica.

The MSEs and average $R^2$ values are computed for each of the estimation methods.

@@

A plot of the MSEs is shown below.

@@

A plot of the $R^2$ values is shown below. The $R^2$ values are much higher than

@@

The features chosen are sorted below.

@@

A plot of Fleiss's kappa values is shown below.

@@

The following plots show the features chosen by each estimation method for all of the cases in the test dataset. The plots are created the by order the features are chosen (from first to third).
